WSClean: a new implementation of a fast, generic wide-field imager

### 2014-01-24 articles/2014/Article-WSClean/wsclean.tex
Progress

\section{The WSClean imager implementation} \label{sec:implementation}
We have implemented the $w$-stacking algorithm in the C++ language. Several steps are multi-threaded in this implementation: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and minor Clean iterations (peak-finding and image subtraction). For the latter, intrinsics were used as well. Because of these optimizations, minor Cleaning iterations are so fast that performing the Clark Clean optimization \citep{clark-clean} of considering only a subset of pixels with a reduced PSF, is not so relevant. However, because the PSF varies for imaging with a non-zero w-value, major iterations are still required to fix the errors made by subtracting a constant PSF during minor iterations.

TODO discuss dealing with memory.

@@

\subsection{W-snapshots}
To get optimal imaging performance, the $w$-values can be kept small by phase-changing towards zenith. As shown, all-sky imaging can be quite efficient with this technique. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

@@

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together.

Another method to shift an image, is by using the periodicity of the FFT function. Since sources outside the field of view will be aliased back into the field, the image can be regridded with the alias in the centre. This complicates gridding, because the antialiasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method.

TODO discuss output standard fits file by using CPIXij keyword \citet{wcs-in-fits}.

@@

>>>>

\section{The WSClean imager implementation} \label{sec:implementation}
We have implemented the $w$-stacking algorithm in the C++ language. Several steps are multi-threaded in this implementation: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and minor Clean iterations (peak-finding and image subtraction). For the latter, intrinsics were used as well. Because of these optimizations, minor Cleaning iterations are so fast that performing the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, is not so relevant. However, because the PSF varies for imaging with a non-zero w-value, major iterations are still required to fix the errors from using a constant PSF during minor iterations.

It will not be possible to store all $w$-layers in memory when creating large images or when many $w$-layers are used. For example, our 32-GB test machine can store 227 $w$-layers of $3072\times3072$ in memory. In more demanding imaging configurations, the implementation performs several passes over the measurement set and will grid a subset of $w$-layers in each pass.

@@

\subsection{W-snapshots}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{MWA imaging example of the Vela and Puppis A supernova remnants. Left: normal projection centred on Puppis A, right: zenith projection to reduce $w$-values, but recentred on Puppis A. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
To get optimal imaging performance, the $w$-values can be kept small by phase-changing towards zenith. As shown, all-sky imaging can be quite efficient with this technique. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

@@

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}. A recentred image can be stored in a Flexible Image Transport System (FITS) file, which allows setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. TODO show difference.

Another method to shift an image, is by using the periodicity of the FFT function. Since sources outside the field of view will be aliased back into the field, the image can be transformed to have the alias in the centre. This complicates gridding during imaging, because the antialiasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method.

@@

TODO: mention full-polarization imaging
